Model based and model free reinforcement learning at FERMI FEL

### 2020-11-25 main.tex
- paper changers

\begin{abstract}
In this paper we discuss a model based reinforcement learning approach in comparison to a model free reinforcement learning approach applied at the FERMI FEL system. Both algorithms and approaches are new in this context and the main purpose of this paper is to show how reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems. In terms of sample-complexity the model-based approach is faster, while the final performance of the model free method is superior - the so called asymptotic performance. The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency.
The implementation is available at \cite{Hirlaender2020}.

\end{abstract}
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
\begin{itemize}
	\item Noise benchmarks with pendulum
	\item Benchmark NAF2 with pendulum
	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
	\item Tailored to accelerators - short horizons are standard?
	\item Improvements of the algorithms:
	\begin{itemize}
		\item Adding the noise feature
		\item Jumping feature using SAC
		\item Noise in the policy
		\item discuss different approaches to e.g. dynamic waiting time
	\end{itemize}
\end{itemize}
%\tableofcontents
\section{Introduction and Motivation}
In particle accelerators one main goal is to provide stable and reproducible performance. In order to achieve this, a number of control problems have to be considered simultaneously. Especially if there is no way to model the physics, one might use optimisation techniques as e.g. derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes to restore or maintain the performance. Another promising way is to apply reinforcement learning which shows several advantages over optimisation methods:
\begin{itemize}
    \item It covers a larger class of problems, RL optimises a sequence of decisions.
    \item It memorizes the problem and does not start always from scratch as an DFO.
    \item Existing data can be used.
    \item The underlying structure of the problem might be deduced.
\end{itemize}
On critical aspect using RL is the number of iteration to train a controller.
In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste. Three algorithms were applied successfully showing different advantages, which will e discussed in the following.

% The methods proposed are the iterative Linear Quadratic Regulator (iLQR) - an extension to non-linear systems of the fundamental control theory Linear Quadratic Regulator (LQR) problem -  and the Normalized Advantage Function (NAF) - a continuous variant of the Q-learning algorithm from reinforcement learning field. An exhaustive presentation of both techniques is provided in the following sections.

In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator. 
\section{The physical set-up}
At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation. Nevertheless, in the last years various approaches have been studied to investigate their applicability in tuning operations. 
%%% Last introduction paragraph
The paper is organised as follows. 
\begin{itemize}
    \item Description of the problem set-up at FERMI
    \item Overview of RL
    \item Details of the implementations used in these studies and theoretical concerns
    \item Results 
    \item Summary and outlook
\end{itemize}

\section{Deep Reinforcement learning}
Assume states $\bs$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
Goal of reinforcement learning is to find a $\policy: \bs\mapsto \ba$, which is the solution of:
\begin{align}
\max J (\policy)  =  \mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],\label{eq:cumulative_reward}
\end{align}
where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, \ba_1,\dots \bs_\horizon,\ba_\horizon)$ drawn by $\policy$ with a horizon $\horizon$.
In the modern field of RL one distinguishes if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g. a deep neural network. In the first case one speaks about policy gradient methods in the latter about approximate dynamic programming, which we now discuss.
\subsection{Approximate dynamic programming}
The state-value function $Q(\bs,\ba)$ is expressed as $Q_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator. By satisfying the Bellmann-optimality equation $Q_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$:
\begin{align}
    \min_\theta \left(\vec Q_\theta - \bellman^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
\end{align}
And $\policy$ can be calculated via:
\begin{align}
	\policy_\theta(\bs)=(\delta(\ba)-\arg\!\max_\ba Q_\theta(\bs,\ba)).
\end{align}
The Bellman operator $\bellman^*$ has a unique fixed point but is non-linear, due to the $\max$ - operator:
\begin{align}
 \bellman^*\vec Q_\theta(\bs_t,\ba_t) := r(\bs_t,\ba_t)+\gamma\max_\ba\left( Q_\theta(\bs_{t+1},\ba) - Q_\theta(\bs_t,\ba_t)\right).
\end{align}
The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{}.
One way to reduce the effect is to take a simple analytical form of the $Q$-function.
\subsection{Normalized advantage function}
If a specific quadratic form of the $Q$ function is assumed:
\begin{align}
     Q_\theta(\bs,\ba) = -\frac{1}{2}(\ba-\mu_\theta(\bs))P_\theta(\ba-\mu_\theta(\bs))^T+V_\theta(\bs).\label{eq:state-action-value-approxiation}
\end{align}

One modification, which is used in these tests is a twin network (weights for network $i$ denoted by $\theta^i$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation. It is motivated by double Q-learning \cite{NIPS2010_091d584f}.
The maximum is given analytically as $\max_\ba Q(\bs,\ba) = V(\bs)$, hence from \cref{eq:minimize_bellmann_optimality}:
\begin{align}
	\min_\theta\left( (\reward(\bs_t,\ba_t)+\gamma \min_{1,2} V_{\theta^i_\text{targ}}(\bs_{t+1}) - (1+\gamma) Q_\theta(\bs_t,\ba_t)\right)^2
\end{align}
$\theta_\text{targ}$ are the weights of a target network, which is softly updated. To stabilize the network training a small artificial noise is added to the actions in the update. This algorithm has an extremely good sample-efficiency for suitable problems as can be found often in accelerators and is used as the baseline for the considered control problem, as it shows very good results.
%\subsection{New NAF network architecture}

\begin{algorithm}[ht]
\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
\begin{algorithmic}[1]
\State initialise $\theta_0$
\For{iteration $k \in [0, \dots, K]$}
\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
%\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
%\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
%\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
%\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
%\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
%\EndFor
\end{algorithmic}
\end{algorithm}

@@

\subsection{Operational deployment}
\NB{Niky} % come abbiamo interagito con FERMI
\subsection{\NB{short FERMI introduction}}
The Free Electron laser Radiation for Multidisciplinary Investigation  better known as the FERMI, is the seeded free-electron laser facility next to the third-generation synchrotron facility ELETTRA at the Elettra Sinchrotrone Trieste laboratories.
A free-electron laser is a fourth-generation light source where the lasing medium consists of very-high-speed electron moving freely through a magnetic structure. The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.

The external optical laser signal provide is contribution to the FEL process in the modulator where it interacts with the relativistic electron beam modulating it in energy. The modulation in energy is the converted into a charge modulation in the dispersive section, and finally the density modulated beams radiation is amplified in the radiators section. The importance of ensuring the best possible overlapping between the seed laser and the electron beam in the modulator is therefore evident.

For this reason the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.

\subsection{\NB{our system: the modulator and the seed laser}}
A more detailed description of the alignment process in the modulator is here provided.
The most critical parameters in a seeded free-electron laser are the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
The problem is simplified by keeping constant the trajectory of the electron beam and the mechanical delay line that controls the temporal alignment.
The final problem faced consists in optimising the seed laser trajectory to match the electron beam and consequently increase the intensity of the FEL radiation.

\section{Proof-of-principle application of RL Agent at FERMI trajectory correction}
\NB{Niky}

@@

\Cref{fig:NAF_training} displays the results, averaged over the two test. A training of 100 episodes (shades area) is followed by a verification of 50 steps. in the upper figure the number of iterations per episode is plotted including the cumulative number of steps.
In the verification phase both algorithms show a similar performance, while the double network needs less training steps, and reveals a more stable overall performance.\\

@@

	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_episodes.pdf}
	\caption{The training on the \emph{NAF2} in the FERMI FEL.}

@@

	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
	\caption{The training .}
	\label{fig:NAF_convergence}

\begin{figure}[!h]
	\centering
	\input{Figures/SL_Alignment_Scheme}
	\caption{The training .}
	\label{fig:comparsion_per}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
	\caption{The training .}
	\label{fig:AE-DYNA_observables}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_verification.pdf}
	\caption{The training .}
	\label{fig:AE-DYNA_verification}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
	\caption{The training .}
	\label{fig:ME-TRPO_observables}
\end{figure}
\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_verification.pdf}
	\caption{The training .}
	\label{fig:ME-TRPO_verification}
\end{figure}

@@

\subsection{NAF2 details}\label{appendix:naf2}
%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
 We compare the different \emph{NAF} variants: \emph{smoothing-double}, \emph{smoothing single}, \emph{standart}. The smoothing adds a small clipped noise on the actions to stabilize the network training as:
 \begin{equation}
 	a'(s') = \text{clip}\left(\mu_{\theta_{\text{targ}}}(s') + \text{clip}(\epsilon,-c,c), a_{Low}, a_{High}\right),
 \end{equation}
 	where $\epsilon \sim \mathcal{N}(0, \sigma)$. This method was used already in \cite{fujimoto2018addressing} to improve the deterministic policy gradient \cite{Silver2014}.
 The double network was used in \cite{fujimoto2018addressing,Haarnoja2018a} and is done in the following way:
 \begin{align}
y(r,s',d) = r + \gamma (1 - d) \min_{i=1,2} Q_{\phi_{i, \text{targ}}}(s', a'(s'))
 \end{align}
and then both are learned by regressing to this target:
\begin{align}
	L(\phi_1, {\mathcal D}) = \mathbb E_{(s,a,r,s',d) \sim {\mathcal D}}{
		\Bigg( Q_{\phi_1}(s,a) - y(r,s',d) \Bigg)^2
	}
\end{align}
and the policy is obtained via:
\begin{align}
\max_{\theta} \underset{s \sim {\mathcal D}}{{\mathrm E}}\left[ Q_{\phi_1}(s, \mu_{\theta}(s)) \right].
\end{align}
The results are shown in \cref{fig:comparsion_smoothing}. One sees the cumulative reward per episode for a training of a total of 50 episodes. The curve labelled \emph{clipping} corresponds to the double network including the smoothing method and shows the best stability during the training yielding quickly a high reward. Also the single network shows good and comparable performance, except for the stability. The worst performance is achieved without smoothing and a single network, nevertheless the result is competing with state of the art model free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_smoothing}
	\caption{Cumulative reward of different NAF implementations as discussed in the text.}
	\label{fig:comparsion_smoothing}
\end{figure}
\subsection{The impact of noise}
A test adding artificial Gaussian noise $\epsilon \sim \mathcal{N}(0, \sigma)$ with $\sigma=0.05$ in the normalized observation space on the states is presented in \cref{fig:comparsion_noise}. There the difference of the three methods becomes even more evident.

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise}
	\caption{Cumulative reward of different NAF implementations with artificial noise as discussed in the text.}
	\label{fig:comparsion_noise}
\end{figure}

>>>>

\begin{abstract}
In this paper we discuss a model based reinforcement learning approach in comparison to a model free reinforcement learning approach applied at the FERMI FEL system. Both algorithms and approaches are new in this context and the main purpose of this paper is to show how reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems. In terms of sample-complexity the model-based approach is faster, while the final performance of the model free method is superior - the so called asymptotic performance. The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency.
%The implementation is available at \cite{Hirlaender2020}.

\end{abstract}
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
\begin{itemize}
	\item Noise benchmarks with pendulum - done
	\item Benchmark NAF2 with pendulum - done
	\item Compare NAF2 and AE-DYNA
	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
	\item Tailored to accelerators - short horizons are standard?
	\item Long waiting time in MBRL
	\item Noise sensivity as mentioned in Benchmark paper
	\item Improvements of the algorithms:
	\begin{itemize}
		\item Adding the noise feature
		\item Jumping feature using SAC
		\item Noise in the policy
		\item discuss different approaches to e.g. dynamic waiting time
		\item usually prior not discussed in papers
	\end{itemize}
\end{itemize}
%\tableofcontents
\section{Introduction and Motivation}
In particle accelerators one main goal is to provide stable and reproducible performance. In order to achieve this, a number of control problems have to be considered simultaneously. Especially if there is no way to model the physics, one might use optimisation techniques as e.g. derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes to restore or maintain the performance. Another promising way is to apply reinforcement learning which shows several advantages over optimisation methods:
\begin{itemize}
    \item It covers a larger class of problems, RL optimises a sequence of decisions.
    \item It memorizes the problem and does not start always from scratch as an DFO.
    \item Existing data can be used.
    \item The underlying structure of the problem might be deduced.
\end{itemize}
One critical aspect using RL is the number of iterations needed to train a controller and the a second critical aspect is the robustness of the training itself.\\
In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste. Three different algorithms were applied successfully to the FERMI FEL problem and reveal different advantages. The critical aspects are addressed and tailored solutions, generally applicable to accelerator control problems are presented.
The paper is organised as follows:
\begin{itemize}
	\item Description of the problem set-up at FERMI
	\item Overview of RL
	\item Details of the implementations used in these studies and theoretical concerns
	\item Results 
	\item Summary and outlook
\end{itemize}
% The methods proposed are the iterative Linear Quadratic Regulator (iLQR) - an extension to non-linear systems of the fundamental control theory Linear Quadratic Regulator (LQR) problem -  and the Normalized Advantage Function (NAF) - a continuous variant of the Q-learning algorithm from reinforcement learning field. An exhaustive presentation of both techniques is provided in the following sections.

\section{The physical set-up of the studied problem}
\begin{figure}[!h]
	\centering
	\input{Figures/SL_Alignment_Scheme}
	\caption{A schematic view on the set-up of the FERMI FEL.}
	\label{fig:schematic_FEL}
\end{figure}
In a seeded free-electron laser one of the most critical parameters is the temporal and transverse overlap of the electron and laser beam in the magnetic section called modulator undulator.\\
At FERMI several beam-based feedback systems are deployed to control the beams trajectories shot to shot with the main purpose of guaranteeing a steady intensity of the light radiation. Nevertheless, in the last years various approaches have been studied to investigate their applicability in tuning operations. \\
A free-electron laser is a fourth-generation light source where the lasing medium consists of very-high-speed electron moving freely through a magnetic structure. The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.\\
The external optical laser signal provide is contribution to the FEL process in the modulator where it interacts with the relativistic electron beam modulating it in energy. The modulation in energy is the converted into a charge modulation in the dispersive section, and finally the density modulated beams radiation is amplified in the radiators section. The importance of ensuring the best possible overlapping between the seed laser and the electron beam in the modulator is therefore evident.\\
For this reason the proposed study focuses on the control of the seed laser trajectory in the modulator, looking at FERMI performance as a reference.
\subsection{\NB{our system: the modulator and the seed laser}}

The most critical parameters in a seeded free-electron laser are the temporal and transverse overlap of the electron and laser beams in the modulator magnetic section. 
The problem is simplified by keeping constant the trajectory of the electron beam and the mechanical delay line that controls the temporal alignment.
A schematic overview of the set-up is provided in \cref{fig:schematic_FEL}.
Two mirrors, TT1 and TT2, are used to control the trajectory of the laser by tilting and  inclining, which gives a total of four degrees of freedom (DOF). In turn the laser overlaps with the electron beam between the two screens, CCD1 and CCD2. Lastly the monitor measures the intensity $I_0$.
The final problem faced consists in optimising the seed laser trajectory to match the electron beam and consequently increase the intensity of the FEL radiation.

%\section{Proof-of-principle application of RL Agent at FERMI trajectory correction}
%\NB{Niky}

\section{Deep Reinforcement learning}
Assume states $\bs$ the set of all states $\states$ and actions $\ba$ all actions $\actions$ and rewards $\reward$ and the set of all rewards $\mathcal{R}$ an initial state distribution $\initstate$.
Goal of reinforcement learning is to find a $\policy^*: \bs\mapsto \ba$, which is the solution of:
\begin{align}
\policy^*\arg\!\max J (\policy)  = 
\mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],\label{eq:cumulative_reward}
\end{align}
where $p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)$ is the distribution of all trajectories $\traj := (\bs_0, \ba_0, \bs_1, \ba_1,\dots \bs_\horizon,\ba_\horizon)$ drawn by $\policy$ with a horizon $\horizon$. $\transitions(\bs_{t+1}|\bs_t,\ba_t)$ characterises the probability to end in a state $\bs_{t+1}$ if an action $\ba_t$ is taken at state $\bs_t$.
In the modern field of RL one distinguishes if the policy $\policy(\ba)\approx \policy_\phi(\bs)$ or the state-action value function $Q(\bs,\ba)$ is approximated using a high capacity function approximator, as e.g. a deep neural network. In the first case $\policy$ is optimized directly and one speaks about policy gradient methods \cite{RichardS.UniversityofAlbertaSutton2018, Williams1992,Baxter2011,pmlr-v28-levine13, Schulman2015,Schulman2017}. In the latter about approximate dynamic programming, which we now discuss.
\section{Model free reinforcement learning - MFRL}\label{s:Model free reinforcement learning}
We do not provide an exhaustive treatment of state of the art MFRL algorithms. Only the main principles are covered for the later statements and an overview is given e.g. in \cite{RichardS.UniversityofAlbertaSutton2018,Levine2020}. We discuss in detail the normalised advantage function algorithm, which has good characteristics for accelerator control problems as shown in \cref{KAIN}.
\subsection{Approximate dynamic programming}
The state-value function $Q(\bs,\ba)$:
\begin{align}
	Q^\policy(\bs,\ba) := \mathbb E_{\traj \sim p_{\policy}(\traj|\bs_t,\ba_t)}\left[\sum_{t=t'}^{\horizon}\discount^{(t'-t)} r(\bs_t,\ba_t)\right],
\end{align}
 is expressed as $Q^\policy_\theta(\bs,\ba)$, where $\theta$ denotes the parameters of the function approximator. By satisfying the Bellmann-optimality equation $Q^\policy_\theta$ can be trained towards the optimal $Q^*(\bs,\ba)$:
\begin{align}
    \min_\theta \left(\vec Q_\theta - \bellman^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
\end{align}
And $\policy$ can be calculated via:
\begin{align}
	\policy_\theta(\bs)=(\delta(\ba)-\arg\!\max_\ba Q_\theta(\bs,\ba)).
\end{align}
The Bellman operator $\bellman^*$ has a unique fixed point but is non-linear, due to the $\max$ - operator:
\begin{align}
 \bellman^*\vec Q_\theta(\bs_t,\ba_t) := r(\bs_t,\ba_t)+\gamma\max_\ba\left( Q_\theta(\bs_{t+1},\ba) - Q_\theta(\bs_t,\ba_t)\right).
\end{align}
The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{Hasselt2015,Mnih2013,Lillicrap2015,Gu2016,Wang2015}.
One way to reduce the effect is to take a simple analytical form of the $Q$-function.
\subsection{Normalized advantage function}
If a specific quadratic form of the $Q$ function is assumed \cite{Gu2016}:
\begin{align}
     Q_\theta(\bs,\ba) = -\frac{1}{2}(\ba-\mu_\theta(\bs))P_\theta(\ba-\mu_\theta(\bs))^T+V_\theta(\bs).\label{eq:state-action-value-approxiation}
\end{align}

One modification, which is used in these tests is a twin network (weights for network $i$ denoted by $\theta^i$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation. It is motivated by double Q-learning \cite{NIPS2010_091d584f,Hasselt2015}.
The maximum is given analytically as $\max_\ba Q(\bs,\ba) = V(\bs)$, hence from \cref{eq:minimize_bellmann_optimality}:
\begin{align}
	\min_\theta\left( (\reward(\bs_t,\ba_t)+\gamma \min_{1,2} V_{\theta^i_\text{targ}}(\bs_{t+1}) - (1+\gamma) Q_\theta(\bs_t,\ba_t)\right)^2
\end{align}
$\theta_\text{targ}$ are the weights of a target network, which is softly updated. To stabilize the network training a small artificial noise is added to the actions in the update. This algorithm has an extremely good sample-efficiency for suitable problems as can be found often in accelerators and is used as the baseline for the considered control problem, as it shows very good results. An illustration and additional changes to the original proposal \cite{Gu2016} and a previous implementation used for experiments using a prioritized replay buffer \cite{Hirlaender2020a} are discussed in \cref{appendix:naf2}.
%\subsection{New NAF network architecture}

%\begin{algorithm}[ht]
%\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
%\begin{algorithmic}[1]
%\State initialise $\theta_0$
%\For{iteration $k \in [0, \dots, K]$}
%\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
%%\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
%%\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
%%\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
%%\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
%%\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
%%\EndFor
%\end{algorithmic}
%\end{algorithm}

@@

@@

\Cref{fig:NAF_training} displays the results, averaged over the two test. A training of 100 episodes was done. In the upper figure the number of iterations per episode is plotted including the cumulative number of steps.
In the verification \cref{fig:NAF_verification} both algorithms show a similar performance, while the double network needs less training steps, and reveals a more stable overall performance.\\

@@

	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_training_episodes.pdf}
	\caption{The training of the \emph{NAF2} on the FERMI FEL.}

@@

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_verification_episodes.pdf}
	\caption{The verification of the \emph{NAF2} on the FERMI FEL.}
	\label{fig:NAF_verification}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
	\caption{The verification of the AE-DYNA-SAC on the FERMI FEL.}
	\label{fig:NAF_convergence}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
	\caption{The training of the AE-DYNA-SAC on the FERMI FEL.}
	\label{fig:AE-DYNA_observables}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_verification.pdf}
	\caption{The verification of the AE-DYNA-SAC on the FERMI FEL.}
	\label{fig:AE-DYNA_verification}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
	\caption{The training of the ME-TRPO on the FERMI FEL.}
	\label{fig:ME-TRPO_observables}
\end{figure}
\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_verification.pdf}
	\caption{The verification of the ME-TRPO on the FERMI FEL.}
	\label{fig:ME-TRPO_verification}
\end{figure}

@@

\subsection{NAF2 details}\label{appendix:naf2}
%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
 We compare the different \emph{NAF} variants: \emph{smoothing-double}, \emph{smoothing single}, \emph{standart}. The smoothing adds a small clipped noise on the actions to stabilize the network training as:
 \begin{equation}
 	a'(s') = \text{clip}\left(\mu_{\theta_{\text{targ}}}(s') + \text{clip}(\epsilon,-c,c), a_{Low}, a_{High}\right),
 \end{equation}
 	where $\epsilon \sim \mathcal{N}(0, \sigma)$. This method was used already in \cite{fujimoto2018addressing} to improve the deterministic policy gradient \cite{Silver2014}.
 The double network was used in \cite{fujimoto2018addressing,Haarnoja2018a} and is done in the following way:
 \begin{align}
y(r,s',d) = r + \gamma (1 - d) \min_{i=1,2} Q_{\phi_{i, \text{targ}}}(s', a'(s'))
 \end{align}
and then both are learned by regressing to this target:
\begin{align}
	L(\phi_1, {\mathcal D}) = \mathbb E_{(s,a,r,s',d) \sim {\mathcal D}}{
		\Bigg( Q_{\phi_1}(s,a) - y(r,s',d) \Bigg)^2
	}
\end{align}
and the policy is obtained via:
\begin{align}
\max_{\theta} \underset{s \sim {\mathcal D}}{{\mathrm E}}\left[ Q_{\phi_1}(s, \mu_{\theta}(s)) \right].
\end{align}
The results are shown in \cref{fig:comparsion_smoothing}. One sees the cumulative reward per episode for a training of a total of 50 episodes. The curve labelled \emph{clipping} corresponds to the double network including the smoothing method and shows the best stability during the training yielding quickly a high reward. Also the single network shows good and comparable performance, except for the stability. The worst performance is achieved without smoothing and a single network, nevertheless the result is competing with state of the art model free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_smoothing}
	\caption{Cumulative reward of different NAF implementations as discussed in the text.}
	\label{fig:comparsion_smoothing}
\end{figure}
\subsection{The impact of noise}
A test adding artificial Gaussian noise $\epsilon \sim \mathcal{N}(0, \sigma)$ with $\sigma=0.05$ in the normalized observation space on the states is presented in \cref{fig:comparsion_noise}. There the difference of the three methods becomes even more evident.The results are shown in \cref{fig:comparsion_noise}. After around 65 episodes the single network without smoothing decreases before reaching the final performance at around 95 episodes, while the smoothing prevents this performance drop. 
\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise}
	\caption{Cumulative reward of different NAF implementations with artificial noise as discussed in the text.}
	\label{fig:comparsion_noise}
\end{figure}
Using the anchors in the dynamics model stabilizes the training of the \emph{AE-DYNA} as illustrated in \cref{fig:comparsion_noise_ae_dyna}. It the upper plot the mean cumulative reward on 10 test episodes on the real environment is shown during the training. It should indicate the result if the training is stopped at this training iteration. One cannot observe this during a real training, unless one does costly performance measurements during the training. The blue curve reaches the target much quicker exhibiting less variation.\\
The lower plot of \cref{fig:comparsion_noise_ae_dyna} shows the batch rewards a measured during the data collection, which is observable during the training.
\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_noise_ae_dyna}
	\caption{Cumulative reward of AE-DYNA with artificial noise using the anchor.}
	\label{fig:comparsion_noise_ae_dyna}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/Comparison_NAF_ae_dyna}
	\caption{The comparison of the NAF2 and the AE-DYNA on the noisy pendulum.}
	\label{fig:comparsion_noise_ae_dyna}
\end{figure}
